Smart Spreading Factor Assignment for LoRaWANs

### 2019-02-25 paper.tex
related hoca
radius
reviewer grammer fix

\subsection{Spreading Factor Assignment Issue}
\par Simultaneous different SF transmissions are orthogonal to each other up to some extent. Which means that, a LoRa GW can simultaneously receive multiple transmissions with different SFs. However, simultaneous transmissions with the same SF may not be received by the GW due to collision. For this reason, SF assignment of nodes is crucial for network performance.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates to GWs. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provide LPWAN solutions to different applications and deployments. Class A end nodes generate uplink transmission at any time and only receive a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using a beacon packet transmitted by GWs. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes with low latency downlink communication, which it requires more power consumption. In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.

@@

\par In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF. Their simulation results show that their proposed algorithms give promising results.

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even they are able to communicate with lower SFs. This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, distribution of SF among nodes becomes an important problem. Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions. In multiple GWs scenarios, this approach may increase the collisions with the nodes in other GW's range. Thus, extra care should be taken for nodes in the intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

@@

\begin{table}
\centering
\caption{PDR for lowest SF and smart SF schemes.}
\label{table:prediction_pdr}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
\multirow{12}{*}{\textbf{R (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}

@@

\begin{table}
\centering
\caption{Prediction accuracy for SVM and DTC.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

>>>>

\subsection{Spreading Factor Assignment Issue}
\par Simultaneous different SF transmissions are orthogonal to each other up to some extent. Which means that, a LoRa gateway (GW) can simultaneously receive multiple transmissions with different SFs. However, simultaneous transmissions with the same SF may not be received by the GW due to collision. For this reason, SF assignment of nodes is crucial for network performance.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates to GWs. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provide LPWAN solutions to different applications and deployments. Class A end nodes generate uplink transmission at any time and only receive a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using a beacon packet transmitted by GWs. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes with low latency downlink communication, which requires more power consumption. In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.

@@

\par In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF. Based on the simulation results, it is shown that the proposed algorithms give promising results.

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs. This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, distribution of SF among nodes becomes an important problem. Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions. In multiple GWs scenarios, this approach may increase the collisions with the nodes in other GW's range. Thus, extra care should be taken for nodes in the intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

@@

\begin{table}
\centering
\caption{PDR for lowest SF and smart SF schemes.}
\label{table:prediction_pdr}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
\multirow{12}{*}{\textbf{r (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}

@@

\begin{table}
\centering
\caption{Prediction accuracy for SVM and DTC.}
\label{table:prediction_accuracy}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
\multirow{8}{*}{\textbf{r (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}
